---
title: "linear model from Web"
author: "Sandra Iovino"
date: "Summer 2024"
output:
   html_document:
         toc: true
         toc_depth: 5
        
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read about the data from the website where it is staged.  Then form three regression models; Model1 is SIMS~ARM, Model2 is SIMS~GRIP, and Model3 is SIMS~ARM+GRIP. For each model find a 95% prediction interval of SIMS given a value of 94  for GRIP and 88 for ARM. Compare Model1 with Model3 using anova. Write it up in a markdown document, push the project up to your github account and submit it back to canvas as link. 

# Methods
all variables are numerical, so we use num~num helper file

```{r include=FALSE}
require(tidyverse)
require(tigerstats)
require(rgl)
require(knitr)

```

```{r}
data <- read.csv(file="https://onlinestatbook.com/case_studies_rvls/physical_strength/data.txt",sep="",header=TRUE)  

```

## Model 1 SIM~ARMS 

### Descriptive results -- Scatterplot  

```{r}
plot(SIMS~ARM,data=data)
```

What this scatter plot graph shows Sims are higher when ARM strength increases. 


### Numerical results

```{r}
cor(SIMS~ARM,data=data)
```
The correlation shows that .6860 is a positive, meaning that the higher strength would make the sims higher. Because if correlation were to be negative it would most likely decrease. 

### Inferential  (Build model.1) 
```{r}
model.1 <- lm(SIMS~ARM,data=data)
summary.lm(model.1)
```
#### scatterplot with model fit

```{r}
plot(SIMS~ARM,data=data)
abline(model.1, col='red')
```

The P-values for the intercept and slope are 0.00000000000000022, So we reject the null because it is smaller than 0.05  


#### Predict at target point

```{r}
newdata = data.frame(ARM =88, GRIP=94)
predict(model.1, newdata, interval='pred')
```

The SIMS prediction for ARM=88 is 0.7, the predicting interval is (-1.73, 3.14)




## Model 2 SIM~GRIP

### Descriptive results -- Scatterplot  

```{r}
plot(SIMS~GRIP,data=data)
```


What this scatter plot graph shows is that the more GRIP strength one has would make the simulation increases also.  


### Numerical results 


```{r}
cor(SIMS~GRIP,data=data)
```
This shows that when Grip strength increases the simulation will be at a positive.



### Inferential  (Build model.2)

```{r}
model.2 <- lm(SIMS~GRIP,data=data)
summary.lm(model.2)
```
  
#### scatterplot with model fit

```{r}
plot(SIMS~GRIP,data=data)
abline(model.2, col='red')
```


This p-value for the interceptor and slope are 0.00000000000000022  


#### predict model.2 at target point

```{r}
newdata = data.frame(ARM =88, GRIP=94)
predict(model.2, newdata, interval='pred')
```
  The SIMS prediction for ARM=88 is 0.5, the predicting interval is (-3.10, 2.03)


## Model 3 SIM~ARM+GRIP
```{r}
model.3 = lm(SIMS~ARM+GRIP, data = data)
summary (model.3)
```
  So we reject 0.00000000000000022
  
### Numerical results

```{r}

df= data.frame(SIMS=data$SIMS,Grip=data$GRIP, ARM=data$ARM)
cor(df)
```
Correlation matrix shows that sims correlates with GRIP AND ARM: And GRIP correlates with SIMS

### Numerical  reults (cor)

```{r}
cor(SIMS~ARM+GRIP, data=data)

```
This is the correlation between the sims score and the two variables ARM and GRIP

### Inferential  (Build 2-dimentional model.3)

```{r}
model.3 = lm(SIMS ~ ARM + GRIP, data=data) 
summary.lm(model.3)
```
p-value from the F-statistics is 0.0000000000000002, this means the data fits the model.
Intercept p-value is 0.0000000000000002
Arm p-value is 0.00000001
Grip p-value is 0.00004
So they are all lower than 0.05

p-value 0.0000000000000002 from the model parameters: Intercept, coefficient of arm and coefficient of grip.


#### predict model.3 at target point

Predict interval of SIMS given a value of 94 for grip and 88 for arm


```{r}
newdata = data.frame(GRIP=94, ARM=88)
predSIMS = predict(model.3, newdata, interval='pred')
print(predSIMS)
```
The predicted sims score for grip=94, arm88 is 0.150
the interval prediction are (-2.13, 2.43)



## Comparing nested models ANOVA Test

### Model.1 vs Model.3

```{r}
anova(model.1, model.3)
summary.lm(model.1)
model.1=lm(SIMS~ ARM, data=data)

```
The nova shows that there is a difference in the models because the p-value is 0.000004

The model 2: SIMS~ARM +GRIP is better than model 1 SIMS~ARM becasue the rss 218 and the rss in model 2is 188

### Model.2 vs Model.3

```{r}
model.2=lm(SIMS ~ GRIP, data=data)
anova(model.2, model.3)
summary.lm(model.1)
```
Model SIM~ARM+GRIP is better than the model sims~grip because the rss is lower (188) and the p-value is 0.000000014


## Informally compare Model.1 with model.2

```{r}
summary.lm(model.1)
summary.lm(model.2)
```
Model.1 sims~arm is better than model.2 sims~grip because the residual for model.1 was lower than model.2